---
title: "ASSSIGNMENT2"
author: "LAURA PAAABY"
date: '2022-08-30'
output: html_document
---
Loading in libraries:
```{r}
library(tidyverse)
```


*1) Use R to figure out how many elements in the vector below are greater than 2 and then tell me what their sum (of the larger than 2 elements) is.*

```{r}
rooms <- c(1, 2, 4, 5, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, 9, 3, NA)
```

The amount of elements that greater than 2:
```{r}
length(which(rooms>2))
```

The sum of all elements greater than 2:
```{r}
sum(which(rooms>2))
```


*2) What type of data is in the 'rooms' vector?*
Lets investigate by looking at the class:
```{r}
class(rooms)
```
They appear to be numeric 



3) Submit the following image to Github: Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object (see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section). Take a screenshot of your RStudio interface showing

a) the line of code you used to create the object,
b) the 'interviews' object in the Environment, and
c) the file structure of your R project in the bottom right "Files" pane.

Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS) or equivalent. Place here the URL leading to the screenshot in your repository.



*4) Challenge: If you managed to create your own Danish king dataset, use it. If not, you the one attached to this assignment (it might need to be cleaned up a bit). Load the dataset into R as a tibble. Calculate the mean() and median() duration of rule over time and find the three mondarchs ruling the longest. How many days did they rule (accounting for transition year?)*

```{r}
#loading in the data:
king_data <- read_csv("/Users/laura/Desktop/GITHUB YES /LauraWP/Homework/DanishMonarchs.csv")
king_data <- as_tibble(king_data)
```

Lets first find the duration:
```{r}
king_data <- king_data %>% 
 mutate(duration = year_of_end_reign - year_of_start_reign) %>% 
  filter(duration != "NA")

### making sure its numeric 
class(king_data$duration)

## finding the mean and median:
mean_dur <- mean(king_data$duration)
median_dur <- median(king_data$duration)

mean_dur
median_dur
```

So the mean duration of each monarchs ruling time is *19.33* and the median is *16.5*.


Lets then find the monarchs ruling the longest:
```{r}
pacman::p_load(Rfast)
library('Rfast')

# The King sitting for the longest 
max(king_data$duration)
king_one <- king_data$Monarch[(which.max(king_data$duration))] 


#The King sitting for the second longest 
nth(king_data$duration, 2, descending = T) #gives length  of duration
nth(king_data$duration, 3, descending = T) #gives length of duration
#### apparently both the second and third most ruling were sitting for 43 years

# The index for those kings are:
which(king_data$duration == (nth(king_data$duration, 2, descending = T)))

#The King sitting for the second longest
king_two <- king_data$Monarch[33] 
king_three <- king_data$Monarch[51] 
```

```{r}
## so the three kings are:
king_one
king_two
king_three
```

Lets now figure out for how many days they been there:

**Christian 4. - 60 years**
```{r}
# how many of these are transition years:
C4_year <- 60
numb_trans_year <- 60/4

# total number of days
((C4_year - numb_trans_year)*365) + (numb_trans_year*366)
```
So Chr 4. have ruled for *21915* days. 

**Christian 9. - 43 years**
```{r}
# how many of these are transition years:
C9_year <- 43
numb_trans_year9 <- 43/4

# total number of days
((C9_year - numb_trans_year9)*365) + (numb_trans_year9*366)
```
So Chr 9. have ruled for *15705.75* days. 

**Erik 7. af Pommern. - 43 years**
Since Erik 7. af Pommern and Christian 9. have both ruled for 43 years, Erik must likewise have ruled *15705.75* days. 












