---
title: "tidytime"
author: "LAURA PAAABY"
date: '2022-08-30'
output: html_document
---

```{r, message=FALSE}

library(tidyverse)
```




```{r}
interview <- read.csv("/Users/laura/Desktop/GITHUB YES /LauraWP/Homework/SAFI_clean.csv")
```

 *TASK 1 - PIPE*
Using pipes, subset the interviews data to include interviews where respondents were members of an irrigation association (memb_assoc) and retain only the columns affect_conflicts, liv_count, and no_meals.

```{r}
interview_pipes <- interview %>% 
  filter(memb_assoc == "yes") %>% 
  select(affect_conflicts, liv_count, no_meals)

head(interview_pipes)
```


 *TASK 2 - MUTATE*
Create a new dataframe from the interviews data that meets the following criteria: 
- contains only the village column and a new column called total_meals containing a value that is equal to the total number of meals served in the household per day on average (no_membrs times no_meals). 

Only the rows where total_meals is greater than 20 should be shown in the final dataframe.

```{r}
interview$no_membrs <- as.numeric(interview$no_membrs)
interview$no_meals <- as.numeric(interview$no_meals)

#### lets go 
interview_mutate <- interview %>% 
  mutate(total_meals = no_membrs*no_meals) %>% 
  mutate(total_meals > 20) %>% 
  select(village, total_meals)

head(interview_mutate)
```


*TASK 3: SPLIT-APPLY-COMBINE .... BY SUMMARIZE*
Example of group by and summarize
```{r}
interview %>%
    group_by(village, memb_assoc) %>%
    summarize(mean_no_membrs = mean(no_membrs))


#### making sure there is no NA's
interview %>%
    filter(!is.na(memb_assoc)) %>%
    group_by(village, memb_assoc) %>%
    summarize(mean_no_membrs = mean(no_membrs))
```

**EXERCISE 3.1**
How many households in the survey have an average of two meals per day? Three meals per day? Are there any other numbers of meals represented?

```{r}
### two
interview %>% 
  count(no_meals == 2, sort = TRUE)


### three
interview %>% 
  count(no_meals == 3, sort = TRUE)

### more
interview %>% 
  count(no_meals > 3, sort = TRUE)

```




**EXERCISE 3.2**
Use group_by() and summarize() to find the mean, min, and max number of household members for each village. Also add the number of observations (hint: see ?n).

```{r}
interview %>% 
  group_by(village) %>% 
  summarize(mean_hh = mean(no_membrs),
            min_hh = min(no_membrs),
            max_hh = max(no_membrs), 
            n())
  

?n

```
**EXERCISE 3.3**
What was the largest household interviewed in each month?
```{r}
# if not already included, add month, year, and day columns
library(lubridate) # load lubridate if not already loaded

interview %>% 
  mutate(month = month(interview_date), 
            day = day(interview_date),
           year = year(interview_date)) %>% 
    group_by(month, year)%>%
    summarize(max_mh = max(no_membrs))
```




# PLOT TIME YES
```{r}
library(ggplot2)
interview %>% 
  ggplot(aes(x = village, y = rooms, colour = respondent_wall_type)) + 
  geom_count() +
  labs(x = "Village yessirs", y = "Number of Rooms", title = "EXERCISE 1")

interview %>% 
  ggplot(aes(x = village, y = rooms, colour = respondent_wall_type)) + 
  geom_point() +
  labs(x = "Village yessirs", y = "Number of Rooms", title = "EXERCISE 1")
  
interview %>% 
  ggplot(aes(x = village, y = rooms, colour = respondent_wall_type)) + 
  geom_jitter() +
  labs(x = "Village yessirs", y = "Number of Rooms", title = "EXERCISE 1")
  
```








## SAve GOOD DaTA 

```{r}
## Not run, but can be used to load in data from previous lesson!
interviews_plotting <- interview %>%
  ## pivot wider by items_owned
  separate_rows(items_owned, sep = ";") %>%
  ## if there were no items listed, changing NA to no_listed_items
  replace_na(list(items_owned = "no_listed_items")) %>%
  mutate(items_owned_logical = TRUE) %>%
  pivot_wider(names_from = items_owned, 
              values_from = items_owned_logical, 
              values_fill = list(items_owned_logical = FALSE)) %>%
  ## pivot wider by months_lack_food
  separate_rows(months_lack_food, sep = ";") %>%
  mutate(months_lack_food_logical = TRUE) %>%
  pivot_wider(names_from = months_lack_food, 
              values_from = months_lack_food_logical, 
              values_fill = list(months_lack_food_logical = FALSE)) %>%
  ## add some summary columns
  mutate(number_months_lack_food = rowSums(select(., Jan:May))) %>%
  mutate(number_items = rowSums(select(., bicycle:car)))
```

